---
title: "Data Building"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=F}
library(tidyr)
library(dplyr)
library(knitr)
<<<<<<< HEAD
url.fl =  "/GDfelids"
opts_chunk$set(warning = FALSE, message = FALSE, root.dir = url.fl)
=======
opts_chunk$set(warning = FALSE, message = FALSE)
>>>>>>> ebf65551c38edf2407127bdc51f5068fbfd7b24c
options(width = 100)
setwd(url.fl)
```

Data building script for grouping GD and PD estimates based on

-   K-means clustering (cluster data set)
-   Country localities (country data set)

```{r, include=F}

<<<<<<< HEAD
#source("Scripts/DataCleanup.R")
load("Data/Popgenden.rda")
load("Data/GPall_data.rda")
=======
#source("Rscript/DataCleanup.R")
load("/Data/GPall_data.rda")
# This data file is not available to public
>>>>>>> ebf65551c38edf2407127bdc51f5068fbfd7b24c
# 3 df objects: gen, den and gp_df

```

# Data description

```{r, echo=FALSE, results='asis'}
kable(table(as.factor(gen$species.specific)) / sum(table(as.factor(gen$species.specific))), caption="Loci type in GD data")
```

There are `r length(unique(gen$ref))` population genetic studies and `r length(unique(den$ref))` population density studies in the raw datasets.

```{r, results='asis', echo=FALSE}
t0 <- gp_df %>% group_by(sp, source) %>% dplyr::summarise(N=n_distinct(population), .groups='drop')
kable(t0 %>% pivot_wider(names_from=source, values_from=N), caption="Number of distinct populations for each measure")
```

```{r, results='asis', echo=FALSE}
t1 <- gp_df %>% group_by(sp, source) %>% dplyr::summarise(N=n_distinct(country), .groups='drop')


kable(t1 %>% pivot_wider(names_from=source, values_from=N), caption="Number of distinct countries in each species")
```

# Applying K-means clustering in each species

Example : tigers

```{r eval=FALSE, include=TRUE}
j = "Panthera tigris"
## Subset to sp data -----------------------------------------------------------------------
sp_name <- sort(unique(gp_df$sp))
sp_df.i <- droplevels(gp_df[gp_df$sp %in% j,])
```

1.  Verify each row has long/lat. Remove duplicated coordinates as we need one point per localisation for clustering.

```{r eval=FALSE, include=TRUE}
all(complete.cases(sp_df.i[,c("long","lat")]))

dups <- duplicated(sp_df.i[,c("long","lat","source")])
if (any(dups)) {
  print(paste(length((dups[dups])), "Dups detected"))
  sp_df <- sp_df.i[!dups,]
  sp_df.ii <- sp_df.i[dups,]
  } else sp_df <- sp_df.i


```

```{r eval=FALSE, include=TRUE}
# - initialise column
  sp_df$cluster_den <- NA
  sp_df$cluster_gen <- NA
  sp_df$closest_ID <- NA
  sp_df$distance <- NA
  sp_df$cluster_center <- NA
```

2.  Overlaid biogeographic regions and apply clustering within each biogeo

```{r eval=FALSE, include=TRUE}
# Biogeographical realms from https://www.sciencebase.gov/catalog/item/508fece8e4b0a1b43c29ca22
# biogeo <- readOGR("www_ter_ecos.shp")
points <- sp::SpatialPoints(coords=sp_df[,c("long","lat")], proj4string = sp::CRS("+proj=longlat +datum=WGS84 +no_defs"))
biogeo.point <- raster::extract(x=biogeo[,c("REALM")], y=points)

sp_df <- cbind(sp_df, biogeo.point[,c("REALM")])
nbiogeo <- table(sp_df$REALM)
nbiogeo

kk <- tapply(sp_df$source, sp_df$REALM, n_distinct)
table(sp_df$REALM, sp_df$source)
# Which biogeo has both GD and PD
sp_df <- subset(sp_df, REALM %in% names(which(kk == 2)))
sp_df$REALM <- droplevels(sp_df$REALM)
#subset to biogeo
for (k in names(kk)){
  bg_i <- which(sp_df$REALM == k)

# Check number of pop in each dataset
foo <- table(sp_df$source[bg_i])
print(k)
print(foo)
```

3.  Apply K-means clustering using GD and PD as initial centroids

```{r eval=FALSE, include=TRUE}
if(length(foo) == 0) next
# using Popden coordinates as centers
cls_df <- sp_df[bg_i,]
den_cls <- kmeans(cls_df[,c("long","lat")], centers=cls_df[which(cls_df$source == "Popden"),c("long","lat")], trace=TRUE)

cls_df[,"cluster_den"] <- den_cls$cluster
cls_df$cluster_den <- paste0(cls_df$cluster_den,"_", cls_df$REALM)

# using Popgen coordinates as centers
gen_cls <- kmeans(cls_df[,c("long","lat")], centers=cls_df[which(cls_df$source == "Popgen"),c("long","lat")], trace=TRUE)
cls_df[,"cluster_gen"] <- gen_cls$cluster
cls_df$cluster_gen <- paste0(cls_df$cluster_gen,"_",cls_df$REALM)

sp_df[bg_i, "cluster_den"] <- cls_df$cluster_den
sp_df[bg_i, "cluster_gen"] <- cls_df$cluster_gen
```

4.  Join unassigned data points to the nearest unassigned point

```{r eval=FALSE, include=TRUE}
## based on gen
bg_df <- sp_df[bg_i, ]
# - calculate distances between the nearest points -Popgen <-> Popden, (in meters)
xx <- bg_df[which(bg_df$source == "Popgen"),c("long","lat")]
yy <- bg_df[which(bg_df$source == "Popden"),c("long","lat")]
mat_gen2den <- geosphere::distm(xx, 
                                yy)
# get distance
bg_df[which(bg_df$source == "Popgen"),"distance"] <- apply(mat_gen2den, 1, min)
bg_df[which(bg_df$source == "Popden"),"distance"] <- apply(mat_gen2den, 2, min)

# get id of the closest and store it in bg_df$closest_ID
bg_df$closest_ID[which(bg_df$source == "Popden")] <- bg_df$ID[which(bg_df$source == "Popgen")][apply(mat_gen2den, 2, which.min)]
bg_df$closest_ID[which(bg_df$source == "Popgen")] <- bg_df$ID[which(bg_df$source == "Popden")][apply(mat_gen2den, 1, which.min)]

# - assign to the nearest cluster (not given by kmeans)
# sort df in ascending order of the distance
bg_df <- bg_df[order(bg_df$distance),]
# by row operation
# modify clus_den
for (i in 1:nrow(bg_df)){
    x <- bg_df[i, "ID"]
      # refresh cluster info
    b <- xtabs(~cluster_den+source, data=bg_df)
    nclus.d <- which(rowSums(b > 0) == 2)
    
    clos_id <- bg_df[bg_df$ID == x,"closest_ID"]
        if(!bg_df$cluster_den[bg_df$ID == clos_id] %in% names(nclus.d)){
      
      bg_df$cluster_den[bg_df$ID == x] <- bg_df[bg_df$ID == clos_id,"cluster_den"]
    }
}
# modify clus_gen
for (i in 1:nrow(bg_df)){
    x <- bg_df[i, "ID"]
      # refresh cluster info
    b <- xtabs(~cluster_gen+source, data=bg_df)
    nclus.g <- which(rowSums(b > 0) == 2)
    clos_id <- bg_df[bg_df$ID == x,"closest_ID"]
        if(!bg_df$cluster_gen[bg_df$ID == clos_id] %in% names(nclus.g)){
          bg_df$cluster_gen[bg_df$ID == x] <- bg_df[bg_df$ID == clos_id,"cluster_gen"]
    }
}

sp_df[bg_i,] <- bg_df
}

```

-   Include duplicates (based on long/lat)

```{r eval=FALSE, include=TRUE}
if(any(dups)){
  poo <- merge(sp_df.ii[,c("long","lat","source")], 
               sp_df[,c("long","lat","source","cluster_den","cluster_gen","cluster_center", "closest_ID","distance")], all.x=T, all.y=F)
  poop <- merge(sp_df.ii, unique(poo))
  sp_df <- rbind(sp_df, poop)
}
```

5.  Check if joined data are within species' dispersal distance

```{r}
# Sigma data -----------------------------------------------------------
sigma_df <- read.csv("Data/Sigma_data.csv")
```

Sigma data details:

-   body mass (bm, in kg), averaged male and female body mass from Johnson et al. 2017, and Huaranca et al. 2019 for *Leopardus jacobita*
-   sigma, σ, averaged across all estimates in SECR models
-   radius, 95% home range radius for a half normal detection, $\sqrt{(-log(0.05)*2*σ^2}$
-   home range (hr) calculated as $2*\pi*radius$
-   maximum distance dispersal (DI) , $40*hr^{0.5}$
-   Generation length in days from Pacifici et al. 2013

```{r, eval=F}
sp_df <- merge(sp_df, sigma_df, by="sp", all.x=T)
sp_df$distance.lowerthanDI <- sp_df$distance <= sp_df$DI*1e3
```

```{r include=FALSE}
## Automatic K-means clustering processing for all species

<<<<<<< HEAD
#source("Scripts/DataBuilding_auto.R") to produce GPclus_results.rds
gp_full <- readRDS("Data/GPclus_results.rds")
=======
#source("Scripts/DataBuilding_auto.R")

>>>>>>> ebf65551c38edf2407127bdc51f5068fbfd7b24c
gp_df <- merge(gp_df, gp_full, all.x=T)
gp_df <- merge(gp_df, sigma_df)
# DI is in km, distance is in m
gp_df$distance.lowerthanDI <- gp_df$distance <= gp_df$DI*1e3

# save cluster results
#saveRDS(gp_df, "Data/Cluster_res_biogeo_v5.rds")

```

6.  Merge GD and PD data based on clusters

```{r, eval=F, include=T}
gendata_sp <- subset(gen, sp %in% j)
dendata_sp <- subset(den, sp %in% j)
clus_sp <- subset(sp_df, sp %in% j)
clus_sp <- clus_sp[which(sp_df$distance.lowerthanDI == TRUE),]
    
    # Separate data based on source
    clus_sp_gen <- clus_sp[which(clus_sp$source == "Popgen"),c("country","long","lat","cluster_center","distance.lowerthanDI")]
    clus_sp_gen <- unique(clus_sp_gen)
    
    clus_sp_den <- clus_sp[which(clus_sp$source == "Popden"),c("ref","country","long","lat","cluster_center","distance.lowerthanDI")]
    clus_sp_den <- unique(clus_sp_den)

    x <- dplyr::left_join(gendata_sp,clus_sp_gen, by=c("country","long","lat"))
    if(any(is.na(x$cluster_center))) x <- x[!is.na(x$cluster_center),]
    xl[[j]] <- x
    
    y <- dplyr::left_join(clus_sp_den, dendata_sp)
    if(any(is.na(y$cluster_center))) y <- y[!is.na(y$cluster_center),]
    y$cluster_center <- droplevels(as.factor(y$cluster_center))
    yl[[j]] <- y

# # Calculate the DP average across studies in each cluster_center before merging with the GD
    DP_mean <- aggregate(dp ~ cluster_center, data = y, FUN=mean, na.action=na.omit)
    DP_max <- aggregate(dp ~ cluster_center, data = y, FUN=max, na.action=na.omit)
    DP_sd <- aggregate(dp ~ cluster_center, data = y, FUN=sd, na.action=na.omit)
    DP_n <- aggregate(dp ~ cluster_center, data = y, FUN=n_distinct, na.action=na.omit)
    DP_median <- aggregate(dp ~ cluster_center, data = y, FUN=median, na.action=na.omit)
    
    DP_mean <- data.frame(cluster_center=DP_max[,1], DP_max=DP_max[,2], DP_mean=DP_mean[,2], DP_sd=DP_sd[,2], DP_n=DP_n[,2], DP_median=DP_median[,2])
    
    DP_ref <- tapply(y$ref, y$cluster_center, unique)
    DP_ref <- lapply(DP_ref, paste, collapse=", ")
    DP_ref <- do.call(rbind,DP_ref)
    DP_ref <- data.frame(ref=DP_ref, cluster_center=rownames(DP_ref))
    
    DP_mean <- merge(DP_mean, DP_ref, all.x=T)
    
    DP_pop <- tapply(y$population, y$cluster_center, unique)
    DP_pop <- lapply(DP_pop, paste, collapse=", ")
    DP_pop <- do.call(rbind,DP_pop)
    DP_pop <- data.frame(population=DP_pop, cluster_center=rownames(DP_pop))
    
    
    DP_mean <- merge(DP_mean, DP_pop, all.x=T)
    
    gen_dp <- dplyr::left_join(x, DP_mean, by="cluster_center", suffix=c(".gen",".den"))
```

```{r eval=TRUE, include=F}
## load cluster data (from 5_Kmeans)
gp_df <- readRDS(paste0(url.fl,"/Data/Cluster_res_biogeo_v5.rds"))
clus.res <- gp_df
<<<<<<< HEAD
=======
source("Script/DataBuilding_Merge_clusters.R")
>>>>>>> ebf65551c38edf2407127bdc51f5068fbfd7b24c

# source("Rscript/DataBuilding_Merge_clusters.R")
# load results from DataBuilding_Merge_clusters.R
load(paste0(url.fl,"/Data/Popgenden.rda"))
```

# Visualise the clustering results

```{r eval=TRUE, include=FALSE}
library(ggmap)
mapp <- list()
sp_name <- unique(gp_df$sp)

for (j in sp_name){
  sp_df <- droplevels(gp_df[gp_df$sp %in% j,])
  print(j)
  sp_df <- unique(sp_df)
  
# create blank map focusing at the so data range  
xy <- cbind(sp_df$long,sp_df$lat)
if(length(unique(sp_df$source)) == 1) next
S <- sp::SpatialPoints(xy)
# add extra margins around the map
S <- raster::buffer(S, width=200000)
# # get map
world <- ggmap::get_map(location = sp::bbox(S), zoom = 4, source="google")

b <- xtabs(~cluster_gen+source, data=sp_df, na.action=na.omit)
nclus <- which(rowSums(b > 0) == 2)
x <- table(sp_df$source[sp_df$cluster_all %in% names(nclus)])

fig.1 <- ggmap(world) +
  geom_point(data = sp_df, aes(x = long, y = lat, pch=factor(source)), size = 3) + 
  scale_shape_manual(values=c(16,4)) +
  geom_path(data = sp_df[which(sp_df$cluster_gen %in% names(nclus)),], aes(x = long, y = lat, col=factor(cluster_gen)), size=2) +
  guides(color="none") +
  geom_point(data =sp_df[which(sp_df$cluster_gen %in% names(nclus)),], 
             aes(x = long, y = lat, col=factor(cluster_gen), pch=factor(source)), size = 3) +
  labs(title=paste0("A) GD (n = ", table(sp_df$source)[2], ")"), 
       subtitle=paste("cluster obtained = ", length(nclus)))  + theme(legend.position="none", plot.margin = margin(0,0,0,0, "cm"))+
  theme(plot.title = element_text(size=12))


b <- xtabs(~cluster_den+source, data=sp_df, na.action=na.omit)
nclus <- which(rowSums(b > 0) == 2)


fig.2 <- ggmap(world) +
  geom_point(data =sp_df, aes(x = long, y = lat, pch=factor(source)), size = 3) + 
  geom_path(data = sp_df[which(sp_df$cluster_den %in% names(nclus)),], aes(x = long, y = lat, col=factor(cluster_den)), size=2) +
  guides(color="none") +
  scale_shape_manual(values=c(16,4)) +
  geom_point(data =sp_df[which(sp_df$cluster_den %in% names(nclus)),], 
             aes(x = long, y = lat, col=factor(cluster_den), pch=factor(source)), size = 3) +
  labs(title=paste0("B) PD (n = ", table(sp_df$source)[1], ")"),
       subtitle=paste("cluster obtained = ", length(nclus)))  + theme(legend.position="none", plot.margin = margin(0,0,0,0, "cm")) +
  theme(plot.title = element_text(size=12))


sp_df.1 <- sp_df[which(sp_df$distance.lowerthanDI),]
sp_df.1 <- droplevels(sp_df.1)

b <- xtabs(~cluster_center+source, data=sp_df.1)
nclus <- which(rowSums(b > 0) == 2)

fig.3 <- ggmap(world) +
  geom_point(data =sp_df, aes(x = long, y = lat, pch=factor(source)), size = 3) + 
  geom_path(data = sp_df.1[which(sp_df.1$cluster_center %in% names(nclus)),], aes(x = long, y = lat, col=factor(cluster_center)), size=2) +
  guides(color="none") +
  scale_shape_manual(values=c(16,4)) +
  geom_point(data =sp_df.1[which(sp_df.1$cluster_center %in% names(nclus)),], 
             aes(x = long, y = lat, col=factor(cluster_center), pch=factor(source)), size = 3) +
  labs(title=paste0("C) dispersal distance (km) = ", unique(sp_df$DI) ),
       subtitle=paste("cluster obtained = ", length(nclus)))  + theme(legend.position="none", plot.margin = margin(0,0,0,0, "cm")) +
  theme(plot.title = element_text(size=12))

# Create a text grob
tgrob <- ggpubr::text_grob(paste(j),size = 13, face = "italic")
# Draw the text
plot_0 <- ggpubr::as_ggplot(tgrob) + theme(plot.margin = margin(0,0.5,0,0, "cm"))

#final <- ggpubr::ggarrange(plot_0, NULL, NULL, fig.1, fig.2, fig.3, nrow=2, ncol=3, common.legend=TRUE, legend = "bottom", heights = c(0.1,1))
#final <- ggpubr::ggarrange(fig.1, fig.2, fig.3, nrow=1, ncol=3, common.legend=TRUE, legend = "none") + theme(plot.margin = margin(0,2,0,0, "cm"))

final <- ggpubr::ggarrange(fig.1,fig.2,fig.3, nrow=1, ncol=3) 

mapp[[j]] <- final
}

#saveRDS(mapp, "Clustering_plots.rds")

```

#### By species {.tabset}

```{r, results='asis', output.width='90%', echo = FALSE, fig_crop=FALSE}
for (i in names(mapp)) {
  cat("##### ",i,"\n")
  par(mar=c(5,4,4,2))
  
  print(mapp[[i]] + theme(plot.margin=margin(1,4,0,0, "cm")))
  

  

  cat(' \n \n')
}
```

# Data summary

```{r, include=FALSE}
nrefgen <- tapply(gen_clus$ref.gen, gen_clus$sp, n_distinct)
ncc <- tapply(gen_clus$cluster_center, gen_clus$sp, n_distinct)

ref.den <- strsplit(gen_clus$ref.den, ",")
gen_clus$ref.den2 <- ref.den

ref.df <- data.frame(sp=with(gen_clus, rep(sp, vapply(ref.den2, length, 1L))), cluster_center=with(gen_clus, rep(cluster_center, vapply(ref.den2, length, 1L))))
ref.df$ref.den <- trimws(unlist(ref.den))
nrefden <- tapply(ref.df$ref.den, ref.df$sp, n_distinct)

tabS1 <- cbind.data.frame(nrefgen, nrefden, ncc)
tabS1$sp <- rownames(tabS1)


gd_rec <- table(gen_clus$sp)
clus_n <- tapply(gen_clus$cluster_center, gen_clus$sp,n_distinct)
pd_rec <- aggregate(DP_n ~ sp, data = gen_clus, sum)

boo <- cbind.data.frame(sp=names(gd_rec),gd_rec=as.vector(gd_rec),pd_rec=pd_rec[,2])
tabS1 <- merge(tabS1, boo)

```

```{r, results='asis', echo=FALSE}
kable(tabS1[,c(1,6,3,5,2,4)], caption="Table S1 : Data summarised for each species in cluster data set", col.names=c("Species","Density records","Density studies","Genetic div. records","Genetic div. studies","Number of clusters with density and genetic records"))
```

```{r, echo=F}
nrefgen <- tapply(gen_count$ref, gen_count$sp, n_distinct)
ncc <- tapply(gen_count$country, gen_count$sp, n_distinct)

ref.den <- strsplit(gen_count$DP_ref, ",")
gen_count$ref.den2 <- ref.den

ref.df <- data.frame(sp=with(gen_count, rep(sp, vapply(ref.den2, length, 1L))), country=with(gen_count, rep(country, vapply(ref.den2, length, 1L))))
ref.df$ref.den <- trimws(unlist(ref.den))
nrefden <- tapply(ref.df$ref.den, ref.df$sp, n_distinct)

tabS1.a <- cbind.data.frame(nrefgen, nrefden, ncc)
tabS1.a$sp <- rownames(tabS1.a)

gd_rec <- table(gen_count$sp)
clus_n <- tapply(gen_count$country, gen_count$sp,n_distinct)
pd_rec <- aggregate(DP_n ~ sp, data = unique(gen_count[,c("DP_n","SxC","sp")]), sum)

boo <- cbind.data.frame(sp=names(gd_rec),gd_rec=as.vector(gd_rec),pd_rec=pd_rec[,2])
tabS1.a <- merge(tabS1.a, boo)
```

```{r, include=T, results='asis', echo=F}
kable(tabS1.a[,c(1,6,3,5,2,4)], caption="Table S1 : Data summarised for each species in country data set", col.names=c("Species","Density records","Density studies","Genetic div. records","Genetic div. studies","Number of countries with density and genetic records"))
```


```{r, include=F}
head(tabS1.a)
```


There are `r sum(tabS1.a$pd_rec)` PD and `r sum(tabS1.a$gd_rec)` GD records from a total of `r sum(tabS1.a$nrefgen, tabS1.a$nrefden)` publications, across `r length(unique(tabS1$sp))` and `r length(unique(tabS1.a$sp))` in cluster and country dataset respectively.

# Global map of cluster

(as in Figure 1)

```{r, include=F}
BSdf <- unique(gen_count[,c("BS","sp")])
BSdf$sp <- as.factor(BSdf$sp)
BSdf$sp <- reorder(BSdf$sp, BSdf$BS)

clus.res <- clus.res[clus.res$distance.lowerthanDI,]
clus.res <- clus.res[!is.na(clus.res$cluster_center),]
clus.res$BS <- NULL
clus.res <- left_join(clus.res, BSdf[,c("BS","sp")])
clus.res$sp <- factor(clus.res$sp, levels=unique(as.character(clus.res$sp)))


no_green = c("#00007F", 
             "#0000B2", "#0000E5", "#0019FF", "#004DFF", "#007FFF", "#00B2FF", 
             "#00E5FF", "#7FFFFF", "#FFFFA5", "#FFFF54", 
             "#FFE500", "#FFB300", "#FF7F00", "#FF4C00", "#FF1900", 
             "#B20000", "#660000")

no_green <- setNames(no_green, levels(BSdf$sp))

# # get map
library(ggmap)
world <- map_data("world", wrap=c(-180,180))
world <- world[-which(world$region == "Antarctica"),]
# shade countries
data_country <-  unique(as.character(gen_count$country))
data_country <- gsub(".*Republic of ", "", data_country)
count_data <- world[which(world$region %in% data_country | world$subregion %in% data_country),]

setdiff(data_country, count_data$region)

basemap <- ggplot(world, aes(x = long, y = lat, group = group)) +
  # geom_tile(data = gplot_wrld_r, 
  #                     aes(x = x, y = y, fill = value), inherit.aes = F, show.legend=F) + 
  # scale_fill_gradient(low = "black", high = "white", na.value = "transparent") +
  geom_polygon(fill = "transparent", colour = "darkgray", size=0.1) 


sum.df <- clus.res
sum.df$cc.sp <- paste(sum.df$sp, sum.df$cluster_center)
boo <- xtabs(~cc.sp+source, data=sum.df)

# Number of clusters with at least one density and genetic data
nclus <- which(rowSums(boo > 0) == 2)

sum.df <- sum.df[sum.df$cc.sp %in% names(nclus),]
sum.df$source <- gsub("Popden", "Population density", sum.df$source)
sum.df$source <- gsub("Popgen", "Population genetic", sum.df$source)

# Add records in country dataset
gp_df$SxC <- paste0(gp_df$sp, "_", gp_df$country)
gp_df.1 <- gp_df[gp_df$SxC %in% intersect(gen_count$SxC, gp_df$SxC),]
gp_df.1$source <- gsub("Popden", "Population density", gp_df.1$source)
gp_df.1$source <- gsub("Popgen", "Population genetic", gp_df.1$source)

map_clus <- basemap + geom_point(data=sum.df, aes(x=long, y=lat, group=cc.sp, col=sp, pch=source), size=2, show.legend=T ) +
  geom_path(data = sum.df, aes(x = long, y = lat, col=sp, group=cc.sp), size = 0.5, lineend = "round", show.legend=FALSE) +
  scale_shape_manual(values=c(16,4)) +
  guides(shape = guide_legend(nrow=2, byrow=T)) +
  #scale_colour_manual(values = no_green[names(no_green) %in% gp_df.1$sp]) +
  scale_colour_manual(values = no_green[names(no_green) %in% gp_df.1$sp], guide =
                        guide_legend(override.aes=list(shape=15, size=4), nrow=3,
                                                                     label.theme=element_text(size=9, 
                                                                                              face="italic"))) +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.text.y=element_blank(),axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),legend.position="bottom",
        panel.background=element_blank(),panel.border=element_rect(color = "black",
                                    fill = NA,
                                    size = 1),panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),plot.background=element_blank(),
        #plot.margin = unit(c(-1, -1, 0, -1), "cm"),
        plot.margin = unit(c(-0.5, -1, 0, -1), "cm"),
        legend.title = element_blank(),
        legend.key = element_blank())

map_count <- basemap +
  geom_polygon(data=count_data, fill = "grey", alpha=0.2, colour = "darkgray", size=0) +
  geom_point(data=gp_df.1, aes(x=long, y=lat, col=sp, group=sp, pch=source), size=1, alpha=1)  +
  scale_shape_manual(values=c(16,4), guide=guide_legend(nrow=2, byrow=T, override.aes=list(size=5), label.theme=element_text(size=9))) +
  scale_colour_manual(values = no_green[names(no_green) %in% gp_df.1$sp], guide =
                        guide_legend(override.aes=list(shape=15, size=4), nrow=3,
                                                                     label.theme=element_text(size=9, 
                                                                                              face="italic"))) +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.text.y=element_blank(),axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),legend.position="bottom",
        panel.background=element_rect(color = NA,
                                    fill = "white",
                                    ),
        panel.border=element_rect(color = "black",
                                    fill = NA,
                                    size = 1),panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),plot.background=element_blank(),
        plot.margin = unit(c(-0.5, -1, 0, -1), "cm"),
        legend.title = element_blank(),
        legend.key = element_blank())

```

```{r, echo=F, results='asis', out.width="0.3\\linewidth"}
ggpubr::ggarrange(map_clus, map_count, nrow=2, heights=c(0.8,1.1), labels=c("A)","B)"))
```
# Separate PD and GD in country map_count
```{r}
## PD
PD_plot <- basemap + 
  geom_point(data=gp_df.1[gp_df.1$source == "Population density",], aes(x=long, y=lat, group=sp, col=sp), pch=16, size=2, show.legend=F ) +
  guides(shape = guide_legend(nrow=2, byrow=T)) +
  scale_colour_manual(values = no_green[names(no_green) %in% gp_df.1$sp]) +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.text.y=element_blank(),axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),legend.position="bottom",
        panel.background=element_blank(),panel.border=element_rect(color = "black",
                                    fill = NA,
                                    size = 1),panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),plot.background=element_blank(),
        plot.margin = unit(c(-1, -1, 0, -1), "cm"),
        legend.title = element_blank(),
        legend.key = element_blank())

## GD
GD_plot <- basemap +
  #geom_polygon(data=count_data, fill = "grey", alpha=0.2, colour = "darkgray", size=0) +
  geom_point(data=gp_df.1[gp_df.1$source == "Population genetic",], 
             aes(x=long, y=lat, col=sp, group=sp), pch=4, size=2, alpha=1, show.legend=F) +
  #scale_colour_manual(values = no_green[names(no_green) %in% gp_df.1$sp], guide =
  #                      guide_legend(override.aes=list(shape=15, size=4), nrow=3,
  #                                                                   label.theme=element_text(size=9, 
  #                                                                                      face="italic"))) +
   guides(shape = guide_legend(nrow=2, byrow=T)) +
  scale_colour_manual(values = no_green[names(no_green) %in% gp_df.1$sp]) +
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.text.y=element_blank(),axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),legend.position="bottom",
        panel.background=element_blank(),panel.border=element_rect(color = "black",
                                    fill = NA,
                                    size = 1),panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),plot.background=element_blank(),
        plot.margin = unit(c(-0.5, -1, 0, -1), "cm"),
        legend.title = element_blank(),
        legend.key = element_blank())

```

```{r, echo=F, results='asis', out.width="0.3\\linewidth"}
ggpubr::ggarrange(PD_plot, GD_plot, map_clus, nrow=3, heights=c(1,1,1.5), labels=c("A)","B)", "C)"))
```
